### 2008-12-19 talk/ecoop2009/rainbow.tex
small fixes

XXX the following section needs a rewriting to be much more high-level and to
compare more directly with classical escape analysis

\anto{Maybe we should talk about ``virtual instances'' and not structures,
  considering the context}

Interpreters for dynamic languages typically continuously allocate a lot of small
objects, for example due to boxing. This makes arithmetic operations extremely
inefficient. For this reason, we
implemented a way for the compiler to try to avoid memory allocations in the
residual code as long as possible. The idea is to try to keep new
run-time structures "exploded": instead of a single run-time object allocated on
the heap, the object is "virtualized" as a set
of fresh local variables, one per field. Only when the object can be accessed by from
somewhere else is it actually allocated on the heap. The effect of this is similar to that of
escape analysis \cite{XXX}, which also prevents allocations of objects that can
be proven to not escape a method or set of methods.

It is not always possible to keep structures virtual.  The main
situation in which it needs to be "forced" (i.e. actually allocated at
run-time) is when the pointer escapes to some non-virtual location like
a field of a real heap structure.  Virtual structures still avoid the run-time
 allocation of most short-lived objects, even in non-trivial situations.

In addition to virtual structures, the compiler can also handle virtual
containers, namely lists and dictionaries \footnote{(R)Python's dictionaries
  are equivalent to .NET \lstinline{Hashtable}s}.  If the indexing operations
can be evaluated at compile-time (i.e., if the variables holding the indexes
are green), the compiler internally keeps track of the state of the container
and store the items as local variables.

@@

\begin{figure}[h]
\begin{center}
\begin{lstlisting}[language=Python]
def interp_eval_abs(args):
    v0 = args[0]
    v1 = IntObj(0)
    a, b = v0, v1
    # hint(a, promote_class=True) implemented as follows:
    cls_a = a.__class__
    switch cls_a:
        default: 
            continue_compilation(jitstate, cls_a)
\end{lstlisting}
\caption{Promotion example 1}
\label{fig:tlc-abs-promotion-1}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\begin{lstlisting}[language=Python]
def interp_eval_abs(args):
    v0 = args[0]
    v1 = IntObj(0)
    a, b = v0, v1
    # hint(a, promote_class=True) implemented as follows:
    cls_a = a.__class__
    switch cls_a:
        IntObj:
            # hint(b, promote_class=True) needs no code
            v0 = IntObj(b.value < a.value)
            if v0.value:
                return a
            else:
                v0 = IntObj(0)
                v1 = args[0]
                a, b = v0, v1
                v0 = IntObj(b.value - a.value)
                return v0
        default: 
            continue_compilation(jitstate, cls_a)
\end{lstlisting}
\caption{Promotion example 2}
\label{fig:tlc-abs-promotion-2}
\end{center}
\end{figure}

>>>>

Interpreters for dynamic languages typically continuously allocate a lot of small
objects, for example due to boxing. This makes arithmetic operations extremely
inefficient. For this reason, we
implemented a way for the compiler to try to avoid memory allocations in the
residual code as long as possible. The idea is to try to keep new
run-time instances "exploded": instead of a single run-time object allocated on
the heap, the object is "virtualized" as a set
of fresh local variables, one per field. Only when the object can be accessed by from
somewhere else is it actually allocated on the heap. The effect of this is similar to that of
escape analysis \cite{Blanchet99escapeanalysis}, \cite{Choi99escapeanalysis},
which also prevents allocations of objects that can be proven to not escape a
method or set of methods (the algorithms however are a lot more advanced than
our very simple analysis).

It is not always possible to keep instances virtual.  The main
situation in which it needs to be "forced" (i.e. actually allocated at
run-time) is when the pointer escapes to some non-virtual location like
a field of a real heap structure.  Virtual instances still avoid the run-time
 allocation of most short-lived objects, even in non-trivial situations.

In addition to virtual instances, the compiler can also handle virtual
containers, namely lists and dictionaries \footnote{(R)Python's dictionaries
  are equivalent to .NET \lstinline{Hashtable}s}.  If the indexing operations
can be evaluated at compile-time (i.e., if the variables holding the indexes
are green), the compiler internally keeps track of the state of the container
and store the items as local variables.

@@

\begin{figure}[h]
\begin{center}
\begin{lstlisting}[language=Python]
def interp_eval_abs(args):
    v0 = args[0]
    v1 = IntObj(0)
    a, b = v0, v1
    # hint(a, promote_class=True) implemented as follows:
    cls_a = a.__class__
    switch cls_a:
        default: 
            continue_compilation(jitstate, cls_a)
\end{lstlisting}
\caption{Promotion step 1}
\label{fig:tlc-abs-promotion-1}
\end{center}
\end{figure}

\begin{figure}[h]
\begin{center}
\begin{lstlisting}[language=Python]
def interp_eval_abs(args):
    v0 = args[0]
    v1 = IntObj(0)
    a, b = v0, v1
    # hint(a, promote_class=True) implemented as follows:
    cls_a = a.__class__
    switch cls_a:
        IntObj:
            # hint(b, promote_class=True) needs no code
            v0 = IntObj(b.value < a.value)
            if v0.value:
                return a
            else:
                v0 = IntObj(0)
                v1 = args[0]
                a, b = v0, v1
                v0 = IntObj(b.value - a.value)
                return v0
        default: 
            continue_compilation(jitstate, cls_a)
\end{lstlisting}
\caption{Promotion step 2}
\label{fig:tlc-abs-promotion-2}
\end{center}
\end{figure}
